---
title: "Modeling approaches for cross-sectional integrative data analysis"
subtitle: "Evaluations and recommendations"
author: "Kenneth Tyler Wilcox & Lijuan Wang"
institute: "Department of Psychology, University of Notre Dame"
date: "IMPS, 21 July 2021"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    self_contained: false # if true, fonts will be stored locally
    lib_dir: libs
    chakra: libs/remark-latest.min.js # Use offline
    nature:
      slideNumberFormat: "%current%"
      beforeInit: "macros.js"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
here::i_am("wilcox-wang-imps2021.Rmd")
options(htmltools.dir.version = FALSE) # Default in xaringanthemer template; don't know what this does
knitr::opts_chunk$set(
  fig.width = 8, fig.height = 4.5, fig.retina = 2, # fig.retina = 2 for most Retina displays (https://yihui.org/knitr/options/#plots)
  dpi = 300,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  fig.show = TRUE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(latex2exp)
library(rlang)
library(stringr)
library(purrr)
library(readr)
library(dplyr)
library(ggplot2)
library(xaringanthemer)
library(here)

# Color codes from Nord theme: https://www.nordtheme.com
style_mono_light(
  base_color       = "#2E3440",
  header_font_google  = google_font("Fira Sans", "400"),
  # header_font_google  = google_font("Roboto", "400"),
  header_h1_font_size = "2.25rem",
  header_h2_font_size = "1.5rem",
  header_h3_font_size = "1.25rem",
  # text_font_google    = google_font("Fira Sans", "300", "300i"),
  text_font_google    = google_font("Roboto", "300", "300i"),
  text_font_size      = "1.25rem", 
  code_font_google    = google_font("Fira Code Retina"),
  header_background_auto = TRUE,
  # title_slide_background_image = "figures/",
  title_slide_background_size = "50%",
  title_slide_background_position = "50% 95%",
  title_slide_background_color = "#ECEFF4", # white
  title_slide_text_color = "#2E3440"
)
options(crayon.enabled = TRUE)
```

class: inverse center middle

# What is Integrative Data Analysis?

## Advantages of IDA

## IDA in Practice

---

class: center middle

# IDA

---

# IDA

+ Definition

--

## Advantages

+ point

--

+ point

--

+ point

---

class: inverse center middle

# Five IDA Models

## Aggregated Regression

## Disaggregated Regression

## Study-Specific Coefficients Regression

## Fixed Slope Multilevel Model

## Random Slopes Multilevel Model

---

# Aggregated Regression

--

+ something

---

# Disaggregated Regression

---

# Study-Specific Coefficients Regression

---

# Fixed Slope Multilevel Model

---

# Random Slopes Multilevel Model

---

class: inverse center middle

# Research Questions

## Simulation Design

## RQ1

## RQ2

## RQ3

---

class: inverse center middle

# Recommendations

---

# Recommendations

+ Don't use ALR

+ FE models don't control type I error rate for either fixed effect with heterogeneity

+ RS MLM can yield well-controlled type I error rates for both fixed effects with appropriate DDFM even in small samples

---

# Conclusions

+ Carefully consider the assumptions involved in any IDA model

+ Suggests consideration of methods like MLMs that directly account for this heterogeneity.

---

# References
