---
title: "Modeling approaches for cross-sectional integrative data analysis"
subtitle: "Evaluations and recommendations"
author: "Kenneth Tyler Wilcox & Lijuan Wang"
institute: "Department of Psychology, University of Notre Dame"
date: "IMPS, 21 July 2021"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    self_contained: false # if true, fonts will be stored locally
    lib_dir: libs
    chakra: libs/remark-latest.min.js # Use offline
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
here::i_am("wilcox-wang-imps2021.Rmd")
options(htmltools.dir.version = FALSE) # Default in xaringanthemer template; don't know what this does
knitr::opts_chunk$set(
  fig.width = 8, fig.height = 4.5, fig.retina = 2, # fig.retina = 2 for most Retina displays (https://yihui.org/knitr/options/#plots)
  dpi = 300,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  fig.show = TRUE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(latex2exp)
library(rlang)
library(stringr)
library(purrr)
library(readr)
library(dplyr)
library(ggplot2)
library(mvtnorm)
library(xaringanthemer)
library(here)

# Color codes from Nord theme: https://www.nordtheme.com
style_mono_light(
  base_color       = "#2E3440",
  header_font_google  = google_font("Fira Sans", "400"),
  # header_font_google  = google_font("Roboto", "400"),
  header_h1_font_size = "2.25rem",
  header_h2_font_size = "1.5rem",
  header_h3_font_size = "1.25rem",
  # text_font_google    = google_font("Fira Sans", "300", "300i"),
  text_font_google    = google_font("Roboto", "300", "300i"),
  text_font_size      = "1.25rem", 
  code_font_google    = google_font("Fira Code"),
  header_background_auto = TRUE,
  # title_slide_background_image = "figures/",
  title_slide_background_size = "50%",
  title_slide_background_position = "50% 95%",
  title_slide_background_color = "#ECEFF4", # white
  title_slide_text_color = "#2E3440"
)
options(crayon.enabled = TRUE)
```

# Integrative Data Analysis (IDA)

> Integrative data analysis (IDA) simultaneously analyzes the *participant-level* data from multiple studies (Curran & Hussong, 2009)

.pull-left[

```{r, echo = FALSE, out.width="100%"}
lit_df <- tibble(time = c("1988-1989", "1990-1999", "2000-2009", "2010-2019"),
                 `# Articles` = c(3, 9, 57, 343))
lit_df %>% 
  ggplot(aes(time, `# Articles`, group = 1)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  scale_x_discrete("") +
  theme(panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank()) +
  theme_xaringan(text_font_size = 24)
```
]
.pull-right[
+ 91% used fixed-effects models

+ Minimal disaggregation

+ But what are we estimating and testing?
]

---

# Participant-Level and Study-Level Effects

```{r, echo = FALSE, message = FALSE, cache = TRUE}
source(here("sim-data-functions.R"))

set.seed(333)
nstudy <- 7
lst_sim <- sim_ida(J = nstudy, nj = rep(30, nstudy), gamma_00 = 0, beta_w = 1, beta_b = -1, sigma2_u0 = 1, sigma2_u1 = 0.5, sigma_u01 = 0)
dat_sim <- lst_sim$data

temp <- dat_sim %>% 
  distinct(x, k, xbar, beta_0k, beta_1k, delta_ik) %>% 
  mutate(gamma_w = 1, gamma_b = -1, gamma_00 = 0)
tempbar <- temp %>% 
  group_by(k) %>% 
  summarise(dbar = mean(delta_ik)) %>% 
  right_join(temp, by = "k") %>% 
  distinct(k, xbar, dbar)

gg_ipd <- ggplot() +
  scale_x_continuous("X", labels = NULL, breaks = NULL) +
  scale_y_continuous("Y", labels = NULL, breaks = NULL) +
  theme_bw() +
  theme(panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank()) +
  geom_smooth(data = temp, aes(x = x, y = delta_ik, group = k), method = "lm",
              fill = NA, color = "black", linetype = 1, size = 1) +
  geom_line(aes(x = temp$x, y = 0 + 1 * temp$x),
                   xend = max(temp$x),
                   yend = 0 + 1 * max(temp$x),
               color = "purple", linetype = 4, size = 1)

gg_ipd_ad <- gg_ipd +
  geom_smooth(data = tempbar, aes(x = xbar, y = dbar), method = "lm",
              fill = NA, color = "red", fullrange = FALSE, linetype = 3, size = 1)

gg_ipd_ad_agg <- gg_ipd_ad +
  geom_line(aes(x = temp$x, y = 0.07312 + (-0.10530) * temp$x),
                   xend = max(temp$x),
                   yend = 0.07132 + (-0.10530) * max(temp$x),
               color = "blue", linetype = 2, size = 1)
```

.pull-left[

```{r, echo = FALSE, out.width="150%"}
gg_ipd_ad_agg + theme_xaringan(text_font_size = 24)
```

+ Aggregated effect: $\gamma_A$ (dashed/blue line)

> "An uninterpretable blend" of $\gamma_W$ and $\gamma_B$
]
.pull-right[

+ Average participant-level effect of $X$ on $Y \rightarrow \gamma_W$ (dot-dashed/purple line)

  + Variability of intercepts $\rightarrow$ $\sigma^2_{u_0}$
  
  + Variability of slopes $\rightarrow$ $\sigma^2_{u_1}$

+ Study-level effect of $\bar{X}$ on $\bar{Y} \rightarrow \gamma_B$ (dotted/red line)
]

.footnote[(Raudenbush & Bryk, 2002, p. 138)]

---

class: inverse center middle

# Research Questions

## 1. What Models Can Disaggregate Participant- and Study-Level Effects?

## 2. How Do We Account for Between-Study Heterogeneity?

## 3. What Methods Work in IDA Small Sample Scenarios?

---

# Recommendations

+ Disaggregate participant-level and study-level fixed effects

+ Carefully consider and model sources of between-study heterogeneity

  + Failing to do so can yield incorrect type I error rates for one or both levels of fixed effects

+ With a small number of studies, random-slopes MLM can yield accurate estimates and well-controlled type I error rates for both types of fixed effects
  + Appropriate degrees of freedom methods are critical
      + Kenward-Roger (1997) DF for study-level fixed effect
      + Containment DF for participant-level fixed effect

+ Overall, MLM can be a viable option for IDA with even as few as six studies

---

# Thanks!

.center[
### `r icon::fa("paper-plane")` kwilcox3@nd.edu
### [`r icon::ii("earth")` www.ktylerwilcox.me](https://www.ktylerwilcox.me)
#### https://www.ktylerwilcox.me/slides/2021imps-wilcox-wang-slides.pdf
### Paper:
Wilcox, K. T., & Wang, L. (In press). Modeling approaches for cross-sectional integrative data analysis: Evaluations and recommendations. *Psychological Methods*. https://doi.org/10.1037/met0000397
]

.footnote[DOI is not yet live.]
